Fundamental Problems with Child and Adolescent Depression Treatments

### 2024-04-25 Manuscript_for_rendering_in_word.qmd
Fixed HAMD and CDRS tables

There are two principal treatment modalities for adolescent depression: psychotherapy and medication. Where does one turn to find the evidence that will inform their treatment options? This question is relevant for patients and their carers, for clinicians and for policy makers. However, the question is particularly difficult to answer for adolescent depression, where there are limited data from head-to-head trials of medication and psychotherapy, and where recommendations must therefore be derived from indirect comparisons of treatment efficacy.

Guidelines, such as the ones that the internationally influential UK National Institute of Health and Care Excellence (NICE) have produced for adolescent depression, are a principal source of such information. NICE recommend that youth are first offered psychological therapy (specifically cognitive behaviour therapy and interpersonal therapy) over medication for most presentations of depression [@niceDepressionChildrenYoung2023]. This conclusion is in keeping with two sources of evidence: medication meta-analyses that cast doubt on the efficacy of most antidepressants, with the exception of fluoxetine, for child and adolescent depression [@ciprianiComparativeEfficacyTolerability2016]; and psychotherapy meta-analyses that conclude psychotherapy to be efficacious for child and adolescent depression [@cuijpersEffectsPsychologicalTreatments2021]. However, such conclusions seem at odds with those of a recent network meta-analysis [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments with each other using both direct (head-to-head) and indirect (indirectly treatment A with treatment C, via studies that directly compare treatments A with B and B with C) evidence. Indeed, the network meta-analysis concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions for adolescent depression?

@@

In this paper, we examine randomised controlled trials (RCTs) of psychotherapy and medication for child and adolescent depression in the following ways. First, we compare key baseline characteristics of medication vs psychotherapy trials, specifically the extent to which they are comparable in a) baseline severity of depression; b) percentage females vs males; c) mean age; and d) number of sites involved. Second, we determine the standardised mean differences of psychotherapy controls vs medication controls using random effects metanalysis. Third, we examine the quality of psychotherapy control conditions by scrutinising the extent to which they are matched to the active intervention in ways such as number or frequency of sessions, and therefore, whether they represent fair pairings from which to draw valid efficacy inferences.

@@

This study was registered on the Open Science Framework (OSF); the protocol can be found [here](https://osf.io/bfmc6). In Supplement X we outline deviations from this pre-registered protocol.

@@

@@

For the purposes of metanalysis, it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, an effect typically not reported. To ensure that our results are not biased by misestimation we simulated n = 1000 datasets for different ranges (0.45 to 0.9) of this correlation and used these datasets in the subsequent analyses (please refer to Supplement for full details).

@@

@@

We ran t-tests to compare the active and control arms of psychotherapy trials on key variables of interest regarding the nature and intensity of the interventions. We extracted data pertaining to the number, duration and intensity of sessions, and the total cumulative hours and period of the intervention. Where a range was provided, the maximum was encoded (e.g. if a paper reported that an intervention involved 8-10 sessions lasting 50-60 minutes, we encoded the number and duration of sessions as 10 and 60, respectively). If sessions varied in frequency across an intervention, we calculated an average by dividing total number of sessions by length of intervention period. Similarly, if the length of sessions varied across the course of the intervention, we calculated a weighted average. Phone call, web-chat and online sessions were encoded as sessions, however guided self-help components were not.

@@

Placebo pill was the control condition for all medication trials. In psychotherapy trials, the control arm included 14 WL controls, 25 care as usual and several other conditions such as 4 attention control conditions. All included trials and the types of treatment controls can be found in Supplementary Table 1.

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: tbl-baseline_results
#| tbl-cap: "Sample characteristics at baseline across medication and psychotherapy studies"
#| tbl-cap-location: top

df_overall_baseline_results %>% 
  mutate(category = case_when(
    row_number() %in% 1:9 ~ "Baseline Severity",
    row_number() %in% 10:18 ~ "Percent Female",
    row_number() %in% 19:27 ~ "Age",
    TRUE ~ ""
  )) %>% # grouping rows
  group_by(category) %>%
  as_flextable(hide_grouplabel = TRUE) %>% 
  paginate(group = "category") %>% 
  bold(i = c(1, 2, 5, 8, 11, 12, 15, 18, 21, 22, 25, 28), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  set_table_properties(layout = "autofit") %>%  # Autofit column widths
  bg(i = c(1, 11, 21), bg  = "#F2F2F2", part = "body") %>% # identify category sections
  padding(i = c(3, 4, 6, 7, 9, 10, 13, 14, 16, 17, 19, 20, 23, 24, 26, 27, 29, 30), j = 1, padding.left = 10) %>% # indent some rows
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Sample characteristics at baseline across medication and psychotherapy studies", colwidths = 8) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 1", colwidths = 8) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  # italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  line_spacing(space = .7, part = "all")

@@

We applied metaregression to obtain the SMDs and confidence intervals of each of the four arms.

@@

Similarly, when we examined the data including only those studies that used the HAM-D (see @fig-plot-means-hamd in Supplementary Materials), medication control and psychotherapy control conditions remained significantly different though the small number of medication studies using the HAM-D resulted in less precise metanalytic estimates of within-group changes.

Further, we showed with our simulations that the effects of different values for the pre-post measures has minimal effect on the estimated outcomes; please refer to @fig-stab-sims in the Supplement.

#### Addressing Regression to the Mean

@@

@@

@@

#Rounding
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(across(where(is.numeric) & !matches("p-value"), ~ round(., 2))) %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, "< 0.001", sprintf("%.3f", `p-value`)))
```

@@

@@

Second, psychotherapy controls have moderate effect size (-0.5) whereas pill-placebo has very large effect size (-1.9). Our analysis could be critiqued as it relies on comparing the within arm symptom change of each trial taken out of randomization, which is generally advised against (CITE). This criticism would apply if our aim were to draw inferences about the efficacy of each arm --- in which case preserving the randomization (in order to balance confounders) is critical. However, we note that, our findings are largely in keeping with those of the network metanalysis, which is designed to preserve the randomization structure. More importantly, we do not make the claim with our analyses that these differences are genuinely due to efficacy differences; they may well be because of the fact that the people who attend psychotherapy and medication trials are different and therefore respond differently. In either case (difference in efficacy vs difference in trial participant profile), the vast disparity in the response to control conditions is reason for major concern about our ability to draw inferences concerning the comparison between the two modalities. This is all the more so as clinicians as well as policy makers often resort to effect sizes to summarise findings, with SMD being the one used as per standard. Our findings make obvious that comparing treatment modalities based on the effect size of each modality is misleading.

Moving beyond the comparison of the controls between the two modalities, we examined whether psychotherapy controls are reasonable counterfactuals to receiving the treatment. An optimal control condition is one where the treatment differs (e.g. treatment vs no treatment), but everything else is equal (in technical terms, the ceteris paribus assumption). An obvious disadvantage of psychotherapy trials is that they are typically unblinded (and hard to blind). Yet, our results show that psychotherapy trials are unlikely to fulfill some other very basic conditions of the "all else is equal" assumption. Most (XXX %) of psychotherapy RCT control conditions are either waitlist control or treatment as usual, both of which are very likely to create negative expectations for participants not randomized to treatment; thus, the comparison is not between treatment and no treatment, but rather treatment and the poor luck of being randomized to waiting. But we also find that even psychological controls were unreasonable counterfactuals. In order to test that a psychological treatment is effective per se (e.g. because of the cognitive techniques the therapist deploys) rather than because of generic effects (e.g. pleasant human contact), aspects such as human contact time should be matched. We find that there are vast differences between treatment and control arms in psychotherapy trials (CITE NUMBER) which may lead to an inflation of the true efficacy of psychological treatments.

@@

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of how these studies are appraised and what needs to be done to create fair and realistic comparators. Indeed, investment of effort and funds should be directed into providing rigorous evidence that established depression psychotherapies are more efficacious than fair controls. There are examples of such psychotherapy RCTS, e.g. in social anxiety [@clarkCognitiveTherapyExposure2006a] where such rigor has been applied.

@@

@@

table_1_cdrs_metareg_baseline <- results_cdrs_metareg_baseline[[1]] %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "term") %>% 
  select(-c("term", "type"))

@@

table_2_cdrs_metareg_baseline <- results_cdrs_metareg_baseline[[2]] %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "baseline_smds")

@@

table_1_hamd_metareg_baseline <- results_hamd_metareg_baseline[[1]] %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "term") %>% 
  select(-c("term", "type"))

@@

table_2_hamd_metareg_baseline <- results_hamd_metareg_baseline[[2]] %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "baseline_smds")

@@

table_mean_coefs_from_sim_adjusted_cohens_d <- df_mean_coefs_from_sim_adjusted_cohens_d %>%
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

@@

# # Tables for baseline comparisons
# 
# # Lets make a table that combines across the sensitivity analyses for severity
# 
# df_severity_sens <- rbind(results_met_sev_cbt, results_met_sev_clin, results_met_sev_no_wl)
# 
# # For row names
# df_severity_sens <- df_severity_sens %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
#     row_number() == 4 ~ "Excluding subclinical",
#     row_number() == 7 ~ "Excluding waitlist",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_severity_sens) <- new_column_names
# 
# # Rounding
# df_severity_sens <- df_severity_sens %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3))
# 
# df_severity_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# flextable(df_severity_sens) %>%
#   set_caption(caption = "Table S1: Sensitivity analyses for comparison of baseline severity levels between psychotherapy and medication trials") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
#   hline(part = "header", border = big_border) %>% 
#   vline(j = 1, border = big_border) %>% 
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")
# 
# # Now sex
# 
# df_sex_sens <- rbind(results_met_perc_women_cbt, results_met_perc_women_clin, results_met_perc_women_no_wl) 
# 
# # For row names
# df_sex_sens <- df_sex_sens %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
#     row_number() == 4 ~ "Excluding subclinical",
#     row_number() == 7 ~ "Excluding waitlist",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_sex_sens) <- new_column_names
# 
# # Rounding
# df_sex_sens <- df_sex_sens %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3))
# 
# df_sex_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# flextable(df_sex_sens) %>%
#   set_caption(caption = "Table S2: Sensitivity analyses for comparison of sex composition between psychotherapy and medication trials") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
#   hline(part = "header", border = big_border) %>% 
#   vline(j = 1, border = big_border) %>% 
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")
# 
# # Now age
# 
# df_age_sens <- rbind(results_met_age_cbt, results_met_age_clin, results_met_age_clin) 
# 
# # For row names
# df_age_sens <- df_age_sens %>%
#   mutate(subgroup = case_when(
#     row_number() == 1 ~ "CBT, fluoxetine and escitalopram",
#     row_number() == 4 ~ "Excluding subclinical",
#     row_number() == 7 ~ "Excluding waitlist",
#     TRUE ~ if_else(subgroup == 0, "Medication", "Psychotherapy")
#   )) 
# 
# # Define new column names
# new_column_names <- c(" ", "K", "Mean", "SE", "Lower CI", "Upper CI", "T2", "p-value")
# 
# # Assign new column names to the data frame
# colnames(df_age_sens) <- new_column_names
# 
# # Rounding
# df_age_sens <- df_age_sens %>%
#   mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
#   mutate_at(vars(ncol(.)), ~ round(., 3))
# 
# df_age_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) 
# 
# flextable(df_age_sens) %>%
#   set_caption(caption = "Table S3: Sensitivity analyses for comparison of age between psychotherapy and medication trials") %>%
#   bold(part = "header") %>%  # Bold the header row
#   bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
#   border_outer(border = small_border) %>%  # Add outer borders
#   hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
#   hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
#   hline(part = "header", border = big_border) %>% 
#   vline(j = 1, border = big_border) %>% 
#   set_table_properties(width = .75) %>%  # Set table width
#   bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
#   autofit() %>%  # Autofit column widths
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
#   padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
#   align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")

>>>>

Should a child or an adolescent receive psychotherapy or medication for their depression, and what information should patients, carers and clinicians use to decide?

<!-- There are two principal treatment modalities for adolescent depression: psychotherapy and medication. Where does one turn to find the evidence that will inform their treatment options? This question is relevant for patients and their carers, for clinicians and for policy makers. However, the question is particularly difficult to answer for adolescent depression, where there are limited data from head-to-head trials of medication and psychotherapy, and where recommendations must therefore be derived from indirect comparisons of treatment efficacy. -->

The internationally influential UK National Institute of Health and Care Excellence (NICE) guidelines for adolescent depression recommend that youth are first offered psychological therapy (specifically cognitive behaviour therapy and interpersonal therapy) over medication for most presentations of depression [@niceDepressionChildrenYoung2023]. This conclusion is in keeping with two sources of evidence: medication meta-analyses that cast doubt on the efficacy of most antidepressants, with the exception of fluoxetine, for child and adolescent depression [@ciprianiComparativeEfficacyTolerability2016]; and psychotherapy meta-analyses that conclude psychotherapy to be efficacious for child and adolescent depression [@cuijpersEffectsPsychologicalTreatments2021]. However, such conclusions seem at odds with those of a recent network meta-analysis [@zhouComparativeEfficacyAcceptability2020], an established method of comparing treatments with each other using both direct (head-to-head) and indirect (indirectly treatment A with treatment C, via studies that directly compare treatments A with B and B with C) evidence. Indeed, the network meta-analysis concluded that only fluoxetine alone and fluoxetine administered together with CBT were significantly more effective than medication control (i.e. pill placebo) or psychotherapy controls. Given this confusing evidence base, how should patients, carers, clinicians and policy makers make treatment decisions for adolescent depression?

@@

In this paper, we examine randomised controlled trials (RCTs) of psychotherapy and medication for child and adolescent depression in the following ways. First, we compare key baseline characteristics of medication and psychotherapy trials, specifically the extent to which they are comparable in a) baseline severity of depression; b) percentage females vs males; c) mean age; and d) number of sites involved. Second, we determine the standardised mean differences of psychotherapy controls and medication controls using random effects metanalysis. Third, we examine the quality of psychotherapy control conditions by scrutinising the extent to which they are matched to the active intervention in ways such as number and frequency of sessions, and therefore, whether they represent fair pairings from which to draw valid efficacy inferences.

@@

This study was registered on the Open Science Framework (OSF); the protocol can be found [here](https://osf.io/bfmc6).

@@

We conducted a series of random effects meta-analyses and tested for subgroup differences between psychotherapy and medication trials in sample characteristics including severity of depressive symptoms at baseline, sex and age. Meta-analyses were implemented using R's Meta package.

@@

For the purposes of metanalysis, it is necessary to estimate a standard error of the SMD. This requires a correlation between the pre- and post-measures, a statistic typically not reported. To ensure that our results are not biased by misestimation we simulated n = 1000 datasets for different values (0.45 to 0.9) of this correlation and used these datasets in the subsequent analyses (please refer to Supplement for full details).

@@

Finally, we examined whether differential regression to the mean may account for differences in effect for psychotherapy and medication trials.

@@

We ran t-tests to compare the active and control arms of psychotherapy trials on key variables of interest regarding the nature and intensity of the interventions. We extracted data pertaining to the number, duration and intensity of sessions, and the total cumulative hours and duration of the intervention. Where a range was provided, the maximum was encoded (e.g. if a paper reported that an intervention involved 8-10 sessions lasting 50-60 minutes, we encoded the number and duration of sessions as 10 and 60, respectively). If sessions varied in frequency across an intervention, we calculated an average by dividing total number of sessions by length of intervention period. Similarly, if the length of sessions varied across the course of the intervention, we calculated a weighted average. Phone call, web-chat and online sessions were encoded as sessions, however guided self-help components were not.

@@

Placebo pill was the control condition for all medication trials. In psychotherapy trials, the control arm included 14 WL controls, 25 care as usual and several other conditions such as 4 attention control conditions. All included trials and the types of treatment controls can be found in Table X in the Supplementary Materials.

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
#| label: tbl-baseline_results
#| tbl-cap: "Sample characteristics at baseline across medication and psychotherapy studies"
#| tbl-cap-location: top
knitr::opts_chunk$set(fp_text_default(bold = TRUE))
df_overall_baseline_results %>% 
  mutate(category = case_when(
    row_number() %in% 1:9 ~ "Baseline Severity",
    row_number() %in% 10:18 ~ "Percent Female",
    row_number() %in% 19:27 ~ "Age",
    TRUE ~ ""
  )) %>% # grouping rows
  group_by(category) %>%
  as_flextable(hide_grouplabel = TRUE) %>% 
  paginate(group = "category") %>% 
  bold(i = c(1, 2, 5, 8, 11, 12, 15, 18, 21, 22, 25, 28), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  # set_table_properties(layout = "fixed") %>%  # Autofit column widths
  bg(i = c(1, 11, 21), bg  = "#F2F2F2", part = "body") %>% # identify category sections
  padding(i = c(3, 4, 6, 7, 9, 10, 13, 14, 16, 17, 19, 20, 23, 24, 26, 27, 29, 30), j = 1, padding.left = 10) %>% # indent some rows
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Sample characteristics at baseline across medication and psychotherapy studies", colwidths = 8) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 1", colwidths = 8) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row  
  # italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "all" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  # padding(i = 2, padding.bottom = 15, part = "header") %>% 
  line_spacing(space = .7, part = "all") %>% 
  # set_table_properties(width = 1, layout = "autofit") %>% 
  autofit()

@@

We applied metaregression to obtain the SMDs and confidence intervals of each of the four study arms.

@@

Similarly, when we examined the data including only those studies that used the HAM-D (see @fig-plot-means-hamd in Supplementary Materials), medication control and psychotherapy control conditions remained significantly different though the small number of medication studies using the HAM-D resulted in less precise metanalytic estimates.

Further, we showed with our simulations that the effects of different values for the pre-post measures had minimal effect on the estimated outcomes; please refer to @fig-stab-sims in the Supplement.

##### Addressing Regression to the Mean

@@

results_cdrs_metareg_baseline[[2]] <- left_join(results_cdrs_metareg_baseline[[2]], study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

@@

results_hamd_metareg_baseline[[2]] <- left_join(results_hamd_metareg_baseline[[2]], study_count_hamd_study[study_count_hamd_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

@@

#Rounding
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(across(where(is.numeric) & !matches("p-value"), ~ round(., 2))) %>%
  mutate(`p-value` = ifelse(`p-value` < 0.001, "< 0.001", sprintf("%.3f", `p-value`)))

@@

```{r, echo=FALSE}
# Filter the dataframe
filtered_df <- df_long_for_metan %>%
  filter(psy_or_med == 1, arm_effect_size == "cohens_d_control")

# Count the occurrences of each type
type_counts <- filtered_df %>%
  count(type) %>%
  mutate(total_count = sum(n),
         wl_cau_sum = sum(n[type %in% c("wl", "cau")]),
         wl_cau_percentage = wl_cau_sum / total_count * 100)

@@

Second, psychotherapy controls have moderate effect size (-0.5) whereas pill-placebo has very large effect size (-1.9). Our analysis could be critiqued as it relies on comparing the within arm symptom change of each trial taken out of randomization, which is generally ill advised. This criticism would apply if our aim were to draw inferences about the efficacy of each arm --- in which case preserving the randomization (in order to balance confounders) is critical. However, we note that, our findings are largely in keeping with those of the network metanalysis, which is designed to preserve the randomization structure. More importantly, we do not make the claim with our analyses that these differences are genuinely due to efficacy differences; they may well be because of the fact that the people who attend psychotherapy and medication trials are different and therefore respond differently. In either case (difference in efficacy vs difference in trial participant profile), the vast disparity in the response to control conditions is reason for major concern about our ability to draw inferences concerning the comparison between the two modalities. This is all the more so as clinicians as well as policy makers often resort to effect sizes to summarise findings, with SMD being the one used as per standard. Our findings make obvious that comparing treatment modalities based on the effect size of each modality is misleading.

Moving beyond the comparison of the controls between the two modalities, we examined whether psychotherapy controls are reasonable counterfactuals to receiving the treatment. An optimal control condition is one where the treatment differs (e.g. treatment vs no treatment), but everything else is equal (in technical terms, the ceteris paribus assumption). An obvious disadvantage of psychotherapy trials is that they are typically unblinded (and hard to blind). Yet, our results show that psychotherapy trials are unlikely to fulfill some other very basic conditions of the "all else is equal" assumption. Most (`r round(type_counts$wl_cau_percentage[1], 2)`%) of psychotherapy RCT control conditions are either waitlist control or treatment as usual, both of which are very likely to create negative expectations for participants not randomized to treatment; thus, the comparison is not between treatment and no treatment, but rather treatment and the poor luck of being randomized to waiting. But we also find that even psychological controls were unreasonable counterfactuals. In order to test that a psychological treatment is effective per se (e.g. because of the cognitive techniques the therapist deploys) rather than because of generic effects (e.g. pleasant human contact), aspects such as human contact time should be matched. We find that there are vast differences between treatment and control arms in psychotherapy trials which may lead to an inflation of the true efficacy of psychological treatments.

@@

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of how these studies are appraised and what needs to be done to create fair and realistic comparators. Indeed, investment of effort and funds should be directed into providing rigorous evidence that established depression psychotherapies are more efficacious than fair controls. There are examples of such psychotherapy RCTs, such as @youngRandomizedDepressionPrevention2016 in the current study, where such rigor has been applied.

@@

table_2_cdrs_metareg_baseline <- results_cdrs_metareg_baseline[[2]] %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "baseline_smds") %>% 
  relocate("n", .before = "baseline_smds") %>% 
  relocate("lower_cis_smds", .before = "upper_cis_smds")

new_column_names <- c("Condition", "N", "Baseline Scores", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(table_2_cdrs_metareg_baseline) <- new_column_names

@@

table_1_cdrs_metareg_baseline <- results_cdrs_metareg_baseline[[1]] %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "term") %>% 
  select(-c("term", "type"))

new_column_names <- c("Condition", "Estimate", "SE", "Statistic", "p value")

# Assign new column names to the data frame
colnames(table_1_cdrs_metareg_baseline) <- new_column_names

@@

@@

table_2_hamd_metareg_baseline <- results_hamd_metareg_baseline[[2]] %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "baseline_smds") %>% 
  relocate("n", .before = "baseline_smds") %>% 
  relocate("lower_cis_smds", .before = "upper_cis_smds")

new_column_names <- c("Condition", "N", "Baseline Scores", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(table_2_hamd_metareg_baseline) <- new_column_names

table_1_hamd_metareg_baseline <- results_hamd_metareg_baseline[[1]] %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition))) %>% 
  relocate("condition", .before = "term") %>% 
  select(-c("term", "type"))

new_column_names <- c("Condition", "Estimate", "SE", "Statistic", "p value")

# Assign new column names to the data frame
colnames(table_1_hamd_metareg_baseline) <- new_column_names

@@

@@

table_mean_coefs_from_sim_adjusted_cohens_d <- df_mean_coefs_from_sim_adjusted_cohens_d %>%
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(condition = case_when(
    condition == "medication_control" ~ "Medication Control",
    condition == "medication_active" ~ "Medication Active",
    condition == "psychotherapy_control" ~ "Psychotherapy Control",
    condition == "psychotherapy_active" ~ "Psychotherapy Active",
    TRUE ~ as.character(condition)))

new_column_names <- c("Condition", "Coefficient", "SE", "Lower CI", "Upper CI")

# Assign new column names to the data frame
colnames(table_mean_coefs_from_sim_adjusted_cohens_d) <- new_column_names

@@
